---
format: html
editor: source
---

```{r}
#| include: false
library(hrbrthemes)

library(sva)
library(vegan)
library(fgsea)
library(DESeq2)
library(limma)
library(BioNERO)
library(permutes)
library(tidyverse)
library(factoextra)
library(clusterProfiler)
library(metafor)

library(ggpubr)
library(ggplot2)
library(pheatmap)
library(multcomp)
library(RColorBrewer)
library(ComplexHeatmap)
library(EnhancedVolcano)

library(DOSE)
library(msigdbr)
library(biomaRt)
library(ReactomePA)
library(org.Hs.eg.db)
library(AnnotationDbi)
library(AnnotationHub)
library(STRINGdb)
library(GSVA)
library(ggfortify)
library(GseaVis)
library(rcartocolor)
library(UpSetR)

library(DT)
library(downloadthis)

mypal <- brewer.pal(8, "Set1")

set.seed(666)
```

```{r}
#| include: false
ensbl2geneid <- read.csv("data/ensbl2geneid.tsv", sep = "\t", header = T)

cell_marker_data <- read.csv("data/cell_marker_human.csv", sep = "\t")
cell_names <- unique(cell_marker_data$cell_name)

meta_data <- read.csv("data/metadata.tsv", sep = "\t")
meta_data <- meta_data[meta_data$Ergicity == "DA",]
colnames(meta_data)[1] <- "sampleid"

meta_data$replicate <- factor(meta_data$Tech.rep)
meta_data$place <- factor(meta_data$Cells.Msc.SPb)

meta_data$pict.name <- sub("_1|_2|_3", "", meta_data$pict.name)

meta_data$mutation <- factor(meta_data$pict.name)
meta_data$condition <- factor(meta_data$Health.PD)

meta_data$File <- meta_data$sampleid
meta_data$File <- paste0(meta_data$File, ".counts")

meta_data <- meta_data %>% relocate(File, .after = sampleid)
meta_data <- meta_data[c(1,2,11:14)]


reads.counts <- read.csv("data/counts.tsv", sep = "\t")
reads.counts$sampleid <- sub(".txt", "", reads.counts$sampleid)
reads.counts$sampleid <- sub("_1A", "-1A", reads.counts$sampleid)
colnames(reads.counts)[4] <- "%_filt"
reads.counts <- merge(meta_data, reads.counts, by = 1)[-c(2,4)]

ddsHTSeq <- DESeqDataSetFromHTSeqCount(sampleTable = meta_data,
                                       directory = "data/htseq/",
                                       design= ~ place + replicate + condition)

rownames(ddsHTSeq) <- sapply(str_split(rownames(ddsHTSeq), "\\."), function(x) x[1])
```

```{r}
#| include: false
ranks.df <- read.csv("data/ranks.tsv", sep = "\t")
ranks.lrrk2.df <- read.csv("data/ranks.lrrk2.tsv", sep = "\t")
ranks.Parkin.df <- read.csv("data/ranks.Parkin.tsv", sep = "\t")
ranks.B.df <- read.csv("data/ranks.B.tsv", sep = "\t")

ranks <- ranks.df$ranks
names(ranks) <- ranks.df$gene_id

ranks.lrrk2 <- ranks.lrrk2.df$ranks.lrrk2
names(ranks.lrrk2) <- ranks.lrrk2.df$gene_id

ranks.Parkin <- ranks.Parkin.df$ranks.Parkin
names(ranks.Parkin) <- ranks.Parkin.df$gene_id

ranks.B <- ranks.B.df$ranks.B
names(ranks.B) <- ranks.B.df$gene_id
```

## Gene set enrichment analysis
<p style="text-align: justify;"> 
To identify biological functions and pathways associated with our experimental conditions, we performed Gene Set Enrichment Analysis (GSEA). This method determines whether predefined gene sets exhibit statistically significant, concordant differences in expression between two biological states. Our analysis utilized gene sets from the MsigDB (HALLMARK), Gene Ontology (GO), KEGG, and Reactome databases. The complete GSEA results are presented in Tables 57-70. Cell type-specific enrichment was further investigated using the Cell Marker 2.0 database (Tables 71-76). Consistent with the differential gene expression analysis, we identified both common and unique molecular pathways for each mutation. The intersections of differential molecular pathways are visualized in Figures 58–61. Enrichment plots for pathways classified as unique or common to LRRK2 and Parkin are provided in Figures 62–103. A summary of the biological interpretation is presented in Table 77.
<p>

<p style="text-align: justify;"> 
Both LRRK2 and Parkin mutations shared a consistent pattern characterized by an upregulation of Transmission across Chemical Synapses (Reactome), coupled with a downregulation of cell cycle and DNA repair processes. These included HALLMARK E2F TARGETS (MSigDB), Homology-Directed Repair (HDR) through Homologous Recombination, and Cell Cycle Checkpoints (Reactome). The LRRK2 mutation was uniquely associated with the upregulation of postsynaptic and axonal components, such as Postsynaptic density membrane, Distal axon, and Regulation of synapse organization (all Gene Ontology), as well as Class A/1 (Rhodopsin-like receptors). Conversely, it exhibited a pronounced downregulation of extracellular matrix (ECM)-related processes, including Extracellular matrix organization, Collagen-containing extracellular matrix, and External encapsulating structure (Gene Ontology). A downregulation of developmental pathways like Sensory organ development and Embryonic organ development was also unique to LRRK2. In contrast, the Parkin mutation displayed a unique downregulation of pathways involved in Epigenetic regulation of gene expression, Regulation of endogenous retroelements, and NGF-stimulated transcription (Reactome). A striking antiphase relationship was observed for ribosomal and translational processes. Pathways such as Ribosomal subunit, Cytosolic ribosome, and Cap-dependent Translation Initiation (Gene Ontology/Reactome) were upregulated in Parkin compared to both LRRK2 and healthy controls, but were downregulated in LRRK2 compared to Parkin and healthy controls. This pattern extended to related processes including Eukaryotic Translation Termination, Selenoamino acid metabolism, and the Response to amino acid deficiency.
<p>

### HALLMARK
```{r}
#| include: false
filter_pval <- function(gsea, NES){
      gsea.df <- as.data.frame(gsea)
      gsea.df <- gsea.df[abs(gsea.df$NES) > NES,]
      gsea.df <- gsea.df[order(gsea.df$NES, decreasing = T),]
      rownames(gsea.df) <- 1:nrow(gsea.df)
      gsea.df <- gsea.df[-2]
      
      return(gsea.df)
}

pathwaysH <- msigdbr(db_species = "HS", species = "Homo sapiens", collection = "H")

GSEA.H <- GSEA(geneList = ranks, pvalueCutoff = 0.05, eps = 0.0, seed = 1, TERM2GENE = pathwaysH[c("gs_name", "ensembl_gene")])
GSEA.H.lrrk2 <- GSEA(geneList = ranks.lrrk2, pvalueCutoff = 0.05, eps = 0.0, seed = 1, TERM2GENE = pathwaysH[c("gs_name", "ensembl_gene")])
GSEA.H.parkin <- GSEA(geneList = ranks.Parkin, pvalueCutoff = 0.05, eps = 0.0, seed = 1, TERM2GENE = pathwaysH[c("gs_name", "ensembl_gene")])
GSEA.H.B <- GSEA(geneList = ranks.B, pvalueCutoff = 0.05, eps = 0.0, seed = 1, TERM2GENE = pathwaysH[c("gs_name", "ensembl_gene")])


GSEA.H_filter <- filter_pval(GSEA.H, 1)
GSEA.H.lrrk2_filter <- filter_pval(GSEA.H.lrrk2, 1)
GSEA.H.parkin_filter <- filter_pval(GSEA.H.parkin, 1)
GSEA.H.B_filter <- filter_pval(GSEA.H.B, 1)
```

#### PD vs Healthy
```{r}
#| echo: false
#| message: false
#| warning: false
datatable(GSEA.H_filter[-c(10,11)], 
          caption = "Table 55. Gene Set Enrichment Analysis (GSEA) of MSigDB HALLMARK gene sets (PD vs Healthy).",
          style = "bootstrap5", 
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip',
                         buttons = c('csv', 'excel')))
```

#### LRRK2 vs Healthy
```{r}
#| echo: false
#| message: false
#| warning: false
datatable(GSEA.H.lrrk2_filter[-c(10,11)], 
          caption = "Table 56. Gene Set Enrichment Analysis (GSEA) of MSigDB HALLMARK gene sets (LRRK2 vs Healthy).",
          style = "bootstrap5", 
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip',
                         buttons = c('csv', 'excel')))
```

#### Parkin vs Healthy
```{r}
#| echo: false
#| message: false
#| warning: false
datatable(GSEA.H.parkin_filter[-c(10,11)], 
          caption = "Table 57. Gene Set Enrichment Analysis (GSEA) of MSigDB HALLMARK gene sets (Parkin vs Healthy).",
          style = "bootstrap5", 
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip',
                         buttons = c('csv', 'excel')))
```

#### LRRK2 vs Parkin
```{r}
#| echo: false
#| message: false
#| warning: false
datatable(GSEA.H.B_filter[-c(10,11)], 
          caption = "Table 58. Gene Set Enrichment Analysis (GSEA) of MSigDB HALLMARK gene sets (LRRK2 vs Parkin).",
          style = "bootstrap5", 
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip',
                         buttons = c('csv', 'excel')))
```

### Gene ontology
```{r}
#| echo: false
gse.GO <- gseGO(geneList = ranks, 
                OrgDb = org.Hs.eg.db, 
                ont = "ALL", 
                minGSSize = 10, 
                maxGSSize = 500, 
                eps = 0.0, 
                pvalueCutoff = 0.05,  
                keyType = "ENSEMBL", 
                verbose = FALSE)

gse.GO.df <- as.data.frame(gse.GO)
gse.GO_filter <- filter_pval(gse.GO.df, 1)

gse.GO.lrrk2 <- gseGO(geneList = ranks.lrrk2, 
                OrgDb = org.Hs.eg.db, 
                ont = "ALL", 
                minGSSize = 10, 
                maxGSSize = 500, 
                eps = 0.0, 
                pvalueCutoff = 0.05,  
                keyType = "ENSEMBL", 
                verbose = FALSE)

gse.GO.lrrk2.df <- as.data.frame(gse.GO.lrrk2)
gse.GO.lrrk2_filter <- filter_pval(gse.GO.lrrk2.df, 1)

gse.GO.parkin <- gseGO(geneList = ranks.Parkin, 
                OrgDb = org.Hs.eg.db, 
                ont = "ALL", 
                minGSSize = 10, 
                maxGSSize = 500, 
                eps = 0.0, 
                pvalueCutoff = 0.05,  
                keyType = "ENSEMBL", 
                verbose = FALSE)

gse.GO.parkin.df <- as.data.frame(gse.GO.parkin)
gse.GO.parkin_filter <- filter_pval(gse.GO.parkin.df, 1)

gse.GO.B <- gseGO(geneList = ranks.B, 
                OrgDb = org.Hs.eg.db, 
                ont = "ALL", 
                minGSSize = 10, 
                maxGSSize = 500, 
                eps = 0.0, 
                pvalueCutoff = 0.05,  
                keyType = "ENSEMBL", 
                verbose = FALSE)

gse.GO.B.df <- as.data.frame(gse.GO.B)
gse.GO.B_filter <- filter_pval(gse.GO.B.df, 1)
```

#### PD vs Healthy
```{r}
#| echo: false
datatable(gse.GO_filter[-c(11:12)], 
          caption = "Table 59. GSEA results obtained using GO terms (PD vs Healthy).",
          style = "bootstrap5", 
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip',
                         buttons = c('csv', 'excel')))
```

#### LRRK2 vs Healthy
```{r}
#| echo: false
datatable(gse.GO.lrrk2_filter[-c(11:12)], 
          caption = "Table 60. GSEA results obtained using GO terms (LRRK2 vs Healthy).",
          style = "bootstrap5", 
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip',
                         buttons = c('csv', 'excel')))
```

#### Parkin vs Healthy
```{r}
#| echo: false
datatable(gse.GO.parkin_filter[-c(11:12)], 
          caption = "Table 61. GSEA results obtained using GO terms (Parkin vs Healthy).",
          style = "bootstrap5", 
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip',
                         buttons = c('csv', 'excel')))
```

#### LRRK2 vs Parkin
```{r}
#| echo: false
datatable(gse.GO.B_filter[-c(11:12)], 
          caption = "Table 62. GSEA results obtained using GO terms (LRRK2 vs Parkin).",
          style = "bootstrap5", 
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip',
                         buttons = c('csv', 'excel')))
```

## KEGG
```{r}
#| include: false
ens2ent <- mapIds(org.Hs.eg.db, keys = rownames(ddsHTSeq), keytype="ENSEMBL", column = "ENTREZID")
ens2ent <- as.data.frame(ens2ent)
ens2ent <- cbind(rownames(ens2ent), ens2ent)

ent.df <- as.data.frame(ranks)
ent.df <- merge(cbind(rownames(ent.df), ent.df), ens2ent, by = 1)
ent.df2 <- ent.df[c(3,2)]
ent.df2 <- ent.df2[!is.na(ent.df2$ens2ent),]
ent.df2 <- ent.df2[!duplicated(ent.df2$ens2ent),]

ent.ranks <- ent.df2$ranks
names(ent.ranks) <- ent.df2$ens2ent

ent.ranks <- sort(ent.ranks, decreasing = T)

ent.lrrk2.df <- as.data.frame(ranks.lrrk2)
ent.lrrk2.df <- merge(cbind(rownames(ent.lrrk2.df), ent.lrrk2.df), ens2ent, by = 1)
ent.lrrk2.df2 <- ent.lrrk2.df[c(3,2)]
ent.lrrk2.df2 <- ent.lrrk2.df2[!is.na(ent.lrrk2.df2$ens2ent),]
ent.lrrk2.df2 <- ent.lrrk2.df2[!duplicated(ent.lrrk2.df2$ens2ent),]

ent.lrrk2.ranks <- ent.lrrk2.df2$ranks
names(ent.lrrk2.ranks) <- ent.lrrk2.df2$ens2ent

ent.lrrk2.ranks <- sort(ent.lrrk2.ranks, decreasing = T)

ent.parkin.df <- as.data.frame(ranks.Parkin)
ent.parkin.df <- merge(cbind(rownames(ent.parkin.df), ent.parkin.df), ens2ent, by = 1)
ent.parkin.df2 <- ent.parkin.df[c(3,2)]
ent.parkin.df2 <- ent.parkin.df2[!is.na(ent.parkin.df2$ens2ent),]
ent.parkin.df2 <- ent.parkin.df2[!duplicated(ent.parkin.df2$ens2ent),]

ent.parkin.ranks <- ent.parkin.df2$ranks
names(ent.parkin.ranks) <- ent.parkin.df2$ens2ent

ent.parkin.ranks <- sort(ent.parkin.ranks, decreasing = T)

ent.B.df <- as.data.frame(ranks.B)
ent.B.df <- merge(cbind(rownames(ent.B.df), ent.B.df), ens2ent, by = 1)
ent.B.df2 <- ent.B.df[c(3,2)]
ent.B.df2 <- ent.B.df2[!is.na(ent.B.df2$ens2ent),]
ent.B.df2 <- ent.B.df2[!duplicated(ent.B.df2$ens2ent),]

ent.B.ranks <- ent.B.df2$ranks
names(ent.B.ranks) <- ent.B.df2$ens2ent

ent.B.ranks <- sort(ent.B.ranks, decreasing = T)
```

```{r}
#| echo: false
#| message: false
#| warning: false
## KEGG
gse.kegg <- gseKEGG(geneList = ent.ranks, organism = 'hsa', minGSSize = 20, maxGSSize = 1000, pvalueCutoff = 0.05, eps = 0.0)

gse.kegg.df <- as.data.frame(gse.kegg)
gse.kegg.df <- gse.kegg[order(gse.kegg.df$NES, decreasing = T),]
rownames(gse.kegg.df) <- 1:nrow(gse.kegg.df)

gse.kegg.lrrk2 <- gseKEGG(geneList = ent.lrrk2.ranks, organism = 'hsa', minGSSize = 20, maxGSSize = 1000, pvalueCutoff = 0.05, eps = 0.0)

gse.kegg.lrrk2.df <- as.data.frame(gse.kegg.lrrk2)
gse.kegg.lrrk2.df <- gse.kegg[order(gse.kegg.lrrk2.df$NES, decreasing = T),]
rownames(gse.kegg.lrrk2.df) <- 1:nrow(gse.kegg.lrrk2.df)

gse.kegg.parkin <- gseKEGG(geneList = ent.parkin.ranks, organism = 'hsa', minGSSize = 20, maxGSSize = 1000, pvalueCutoff = 0.05, eps = 0.0)

gse.kegg.parkin.df <- as.data.frame(gse.kegg.parkin)
gse.kegg.parkin.df <- gse.kegg[order(gse.kegg.parkin.df$NES, decreasing = T),]
rownames(gse.kegg.parkin.df) <- 1:nrow(gse.kegg.parkin.df)

gse.kegg.B <- gseKEGG(geneList = ent.B.ranks, organism = 'hsa', minGSSize = 20, maxGSSize = 1000, pvalueCutoff = 0.05, eps = 0.0)

gse.kegg.B.df <- as.data.frame(gse.kegg.B)
gse.kegg.B.df <- gse.kegg[order(gse.kegg.B.df$NES, decreasing = T),]
rownames(gse.kegg.B.df) <- 1:nrow(gse.kegg.B.df)
```

#### PD vs Healthy
```{r}
#| echo: false
datatable(gse.kegg.df[-c(11:12)], 
          caption = "Table 63. GSEA results obtained using KEGG (PD vs Healthy).",
          style = "bootstrap5", 
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip',
                         buttons = c('csv', 'excel')))
```

#### LRRK2 vs Healthy
```{r}
#| echo: false
datatable(gse.kegg.lrrk2.df[-c(11:12)], 
          caption = "Table 64. GSEA results obtained using KEGG (LRRK2 vs Healthy).",
          style = "bootstrap5", 
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip',
                         buttons = c('csv', 'excel')))
```

#### Parkin vs Healthy
```{r}
#| echo: false
datatable(gse.kegg.parkin.df[-c(11:12)], 
          caption = "Table 65. GSEA results obtained using KEGG (Parkin vs Healthy).",
          style = "bootstrap5", 
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip',
                         buttons = c('csv', 'excel')))
```

#### LRRK2 vs Parkin
```{r}
#| echo: false
datatable(gse.kegg.B.df[-c(11:12)], 
          caption = "Table 66. GSEA results obtained using KEGG (LRRK2 vs Parkin).",
          style = "bootstrap5", 
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip',
                         buttons = c('csv', 'excel')))
```

### Reactome
```{r}
#| include: false
gse.Reactome <- gsePathway(ent.ranks, pvalueCutoff = 0.05, eps = 0.0, seed = 666)
gse.Reactome.lrrk2 <- gsePathway(ent.lrrk2.ranks, pvalueCutoff = 0.05, eps = 0.0, seed = 666)
gse.Reactome.parkin <- gsePathway(ent.parkin.ranks, pvalueCutoff = 0.05, eps = 0.0, seed = 666)
gse.Reactome.B <- gsePathway(ent.B.ranks, pvalueCutoff = 0.05, eps = 0.0, seed = 666)

gse.Reactome.df <- as.data.frame(gse.Reactome)
gse.Reactome.df <- gse.Reactome.df[order(gse.Reactome.df$NES, decreasing = T),]
rownames(gse.Reactome.df) <- 1:nrow(gse.Reactome.df)

gse.Reactome.lrrk2.df <- as.data.frame(gse.Reactome.lrrk2)
gse.Reactome.lrrk2.df <- gse.Reactome.lrrk2.df[order(gse.Reactome.lrrk2.df$NES, decreasing = T),]
rownames(gse.Reactome.lrrk2.df) <- 1:nrow(gse.Reactome.lrrk2.df)

gse.Reactome.parkin.df <- as.data.frame(gse.Reactome.parkin)
gse.Reactome.parkin.df <- gse.Reactome.parkin.df[order(gse.Reactome.parkin.df$NES, decreasing = T),]
rownames(gse.Reactome.parkin.df) <- 1:nrow(gse.Reactome.parkin.df)

gse.Reactome.B.df <- as.data.frame(gse.Reactome.B)
gse.Reactome.B.df <- gse.Reactome.B.df[order(gse.Reactome.B.df$NES, decreasing = T),]
rownames(gse.Reactome.B.df) <- 1:nrow(gse.Reactome.B.df)
```

#### PD vs Healthy
```{r}
#| echo: false
datatable(gse.Reactome.df[-c(11:12)], 
          caption = "Table 67. GSEA results obtained using Reactome (PD vs Healthy).",
          style = "bootstrap5", 
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip',
                         buttons = c('csv', 'excel')))
```

#### LRRK2 vs Healthy
```{r}
#| echo: false
datatable(gse.Reactome.lrrk2.df[-c(11:12)], 
          caption = "Table 68. GSEA results obtained using Reactome (LRRK2 vs Healthy).",
          style = "bootstrap5", 
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip',
                         buttons = c('csv', 'excel')))
```

#### Parkin vs Healthy
```{r}
#| echo: false
datatable(gse.Reactome.parkin.df[-c(11:12)], 
          caption = "Table 69. GSEA results obtained using Reactome (Parkin vs Healthy).",
          style = "bootstrap5", 
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip',
                         buttons = c('csv', 'excel')))
```

#### LRRK2 vs Parkin
```{r}
#| echo: false
datatable(gse.Reactome.B.df[-c(11:12)], 
          caption = "Table 70. GSEA results obtained using Reactome (LRRK2 vs Parkin).",
          style = "bootstrap5", 
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip',
                         buttons = c('csv', 'excel')))
```

### Cell Marker 2.0
```{r}
#| message: false
#| warning: false
#| include: false
cell_marker_data <- read.csv("data/cell_marker_human.csv", sep = "\t")
cell_names <- unique(cell_marker_data$cell_name)

cells <- cell_marker_data %>%
    dplyr::select(cell_name, GeneID)

get_GSEA <- function(ranks.df, TERM2GENE, 
                     eps, pvalueCutoff, minGSSize, maxGSSize, pAdjustMethod, NES){
    gsea.df <- GSEA(ranks.df, TERM2GENE = TERM2GENE, eps = eps, 
                    pvalueCutoff = pvalueCutoff, minGSSize = minGSSize, 
                    maxGSSize = maxGSSize, pAdjustMethod = pAdjustMethod)
    gsea.df <- as.data.frame(gsea.df)
    gsea.df <- gsea.df[abs(gsea.df$NES) > NES,]
    gsea.df <- gsea.df[order(gsea.df$NES, decreasing = T),]
    return(gsea.df)
}

gsea.cellmarker <- get_GSEA(ent.ranks, 
                            TERM2GENE = cells, 
                            minGSSize = 10, 
                            maxGSSize = 500, 
                            eps = 0, 
                            pAdjustMethod = "BH", 
                            pvalueCutoff = 0.05, 
                            NES = 2)
gsea.cellmarker.df <- as.data.frame(gsea.cellmarker)
rownames(gsea.cellmarker.df) <- 1:nrow(gsea.cellmarker.df)

gsea.cellmarker.lrrk2 <- get_GSEA(ent.lrrk2.ranks, 
                            TERM2GENE = cells, 
                            minGSSize = 10, 
                            maxGSSize = 500, 
                            eps = 0, 
                            pAdjustMethod = "BH", 
                            pvalueCutoff = 0.05, 
                            NES = 2)
gsea.cellmarker.lrrk2.df <- as.data.frame(gsea.cellmarker.lrrk2)
rownames(gsea.cellmarker.lrrk2.df) <- 1:nrow(gsea.cellmarker.lrrk2.df)

gsea.cellmarker.B <- get_GSEA(ent.B.ranks, 
                            TERM2GENE = cells, 
                            minGSSize = 10, 
                            maxGSSize = 500, 
                            eps = 0, 
                            pAdjustMethod = "BH", 
                            pvalueCutoff = 0.05, 
                            NES = 2)
gsea.cellmarker.B.df <- as.data.frame(gsea.cellmarker.B)
rownames(gsea.cellmarker.B.df) <- 1:nrow(gsea.cellmarker.B.df)
```

#### PD vs Healthy
```{r}
#| echo: false
datatable(gsea.cellmarker.df[-c(11:12)], 
          caption = "Table 71. GSEA results obtained using Cell Marker 2.0 (PD vs Healthy).",
          style = "bootstrap5", 
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip',
                         buttons = c('csv', 'excel')))
```

#### LRRK2 vs Healthy
```{r}
#| echo: false
datatable(gsea.cellmarker.lrrk2.df[-c(11:12)], 
          caption = "Table 72. GSEA results obtained using Cell Marker 2.0 (LRRK2 vs Healthy).",
          style = "bootstrap5", 
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip',
                         buttons = c('csv', 'excel')))
```

#### LRRK2 vs Parkin
```{r}
#| echo: false
datatable(gsea.cellmarker.B.df[-c(11:12)], 
          caption = "Table 73. GSEA results obtained using Cell Marker 2.0 (LRRK2 vs Parkin).",
          style = "bootstrap5", 
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip',
                         buttons = c('csv', 'excel')))
```

#### Only brain cell markers
```{r}
#| include: false
cell_brain <- cell_marker_data[cell_marker_data$tissue_class == "Brain",]
cell_brain <- cell_brain[c("cell_name", "GeneID")]

gsea.cellmarker.brain <- get_GSEA(ent.ranks, 
                                  TERM2GENE = cell_brain, 
                                  minGSSize = 10, 
                                  maxGSSize = 500, 
                                  eps = 0, 
                                  pAdjustMethod = "BH", 
                                  pvalueCutoff = 0.05,
                                  NES = 1)
gsea.cellmarker.brain.df <- as.data.frame(gsea.cellmarker.brain)
rownames(gsea.cellmarker.brain.df) <- 1:nrow(gsea.cellmarker.brain.df)

gsea.cellmarker.brain.lrrk2 <- get_GSEA(ent.lrrk2.ranks, 
                                  TERM2GENE = cell_brain, 
                                  minGSSize = 10, 
                                  maxGSSize = 500, 
                                  eps = 0, 
                                  pAdjustMethod = "BH", 
                                  pvalueCutoff = 0.05,
                                  NES = 1)
gsea.cellmarker.brain.lrrk2.df <- as.data.frame(gsea.cellmarker.brain.lrrk2)
rownames(gsea.cellmarker.brain.lrrk2.df) <- 1:nrow(gsea.cellmarker.brain.lrrk2.df)

gsea.cellmarker.brain.B <- get_GSEA(ent.B.ranks, 
                                  TERM2GENE = cell_brain, 
                                  minGSSize = 10, 
                                  maxGSSize = 500, 
                                  eps = 0, 
                                  pAdjustMethod = "BH", 
                                  pvalueCutoff = 0.05,
                                  NES = 1)
gsea.cellmarker.brain.B.df <- as.data.frame(gsea.cellmarker.brain.B)
rownames(gsea.cellmarker.brain.B.df) <- 1:nrow(gsea.cellmarker.brain.B.df)
```

##### PD vs Healthy
```{r}
#| echo: false
datatable(gsea.cellmarker.brain.df[-c(11:12)], 
          caption = "Table 74. GSEA results obtained using Reactome - only brain cell markers (PD vs Healthy).",
          style = "bootstrap5", 
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip',
                         buttons = c('csv', 'excel')))
```

##### LRRK2 vs Healthy
```{r}
#| echo: false
datatable(gsea.cellmarker.brain.lrrk2.df[-c(11:12)], 
          caption = "Table 75. GSEA results obtained using Reactome - only brain cell markers (LRRK2 vs Healthy).",
          style = "bootstrap5", 
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip',
                         buttons = c('csv', 'excel')))
```

##### LRRK2 vs Parkin
```{r}
#| echo: false
datatable(gsea.cellmarker.brain.B.df[-c(11:12)], 
          caption = "Table 76. GSEA results obtained using Reactome - only brain cell markers (LRRK2 vs Parkin).",
          style = "bootstrap5", 
          extensions = 'Buttons', 
          options = list(dom = 'lBfrtip',
                         buttons = c('csv', 'excel')))
```

## Intersections between pathways
### HALLMARK
```{r}
#| include: false
HALLMARK_list <- list(PD_ALL = GSEA.H_filter$ID[GSEA.H_filter$NES > 0],
                      Healthy_ALL = GSEA.H_filter$ID[GSEA.H_filter$NES < 0],
                      PD_LRKK2 = GSEA.H.lrrk2_filter$ID[GSEA.H.lrrk2_filter$NES > 0],
                      Healthy_LRKK2 = GSEA.H.lrrk2_filter$ID[GSEA.H.lrrk2_filter$NES < 0],
                      PD_Parkin = GSEA.H.parkin_filter$ID[GSEA.H.parkin_filter$NES > 0],
                      Healthy_Parkin = GSEA.H.parkin_filter$ID[GSEA.H.parkin_filter$NES < 0],
                      LRRK2 = GSEA.H.B_filter$ID[GSEA.H.B_filter$NES > 0],
                      Parkin = GSEA.H.B_filter$ID[GSEA.H.B_filter$NES < 0])

HALLMARK_upset <- upset(fromList(HALLMARK_list), 
      nintersects = 40, 
      nsets = 8, 
      order.by = "freq", 
      decreasing = T, 
      mb.ratio = c(0.6, 0.4),
      number.angles = 0, 
      text.scale = 1.1, 
      point.size = 2.8, 
      line.size = 1
)

# pdf("figures/HALLMARK_upset.pdf", width = 9, height = 7)
# HALLMARK_upset
# dev.off()
```

[Figure 58](figures/HALLMARK_upset.pdf)

```{r fig.width=9, fig.height=7}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 58. Intersections between HALLMARK pathways."
HALLMARK_upset
```

### GO
```{r}
#| include: false
GO_list <- list(PD_ALL = gse.GO_filter$Description[gse.GO_filter$NES > 0],
                      Healthy_ALL = gse.GO_filter$Description[gse.GO_filter$NES < 0],
                      PD_LRKK2 = gse.GO.lrrk2_filter$Description[gse.GO.lrrk2_filter$NES > 0],
                      Healthy_LRKK2 = gse.GO.lrrk2_filter$Description[gse.GO.lrrk2_filter$NES < 0],
                      PD_Parkin = gse.GO.parkin_filter$Description[gse.GO.parkin_filter$NES > 0],
                      Healthy_Parkin = gse.GO.parkin_filter$Description[gse.GO.parkin_filter$NES < 0],
                      LRRK2 = gse.GO.B_filter$Description[gse.GO.B_filter$NES > 0],
                      Parkin = gse.GO.B_filter$Description[gse.GO.B_filter$NES < 0])

GO_upset <- upset(fromList(GO_list), 
      nintersects = 40, 
      nsets = 8, 
      order.by = "freq", 
      decreasing = T, 
      mb.ratio = c(0.6, 0.4),
      number.angles = 0, 
      text.scale = 1.1, 
      point.size = 2.8, 
      line.size = 1
)

# pdf("figures/GO_upset.pdf", width = 9, height = 7)
# GO_upset
# dev.off()
```

[Figure 59](figures/GO_upset.pdf)

```{r fig.width=9, fig.height=7}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 59. Intersections between GO pathways."
GO_upset
```

### KEGG
```{r}
#| include: false
KEGG_list <- list(PD_ALL = gse.kegg.df$Description[gse.kegg.df$NES > 0],
                      Healthy_ALL = gse.kegg.df$Description[gse.kegg.df$NES < 0],
                      PD_LRKK2 = gse.kegg.lrrk2.df$Description[gse.kegg.lrrk2.df$NES > 0],
                      Healthy_LRKK2 = gse.kegg.lrrk2.df$Description[gse.kegg.lrrk2.df$NES < 0],
                      PD_Parkin = gse.kegg.parkin.df$Description[gse.kegg.parkin.df$NES > 0],
                      Healthy_Parkin = gse.kegg.parkin.df$Description[gse.kegg.parkin.df$NES < 0],
                      LRRK2 = gse.kegg.B.df$Description[gse.kegg.B.df$NES > 0],
                      Parkin = gse.kegg.B.df$Description[gse.kegg.B.df$NES < 0])

KEGG_upset <- upset(fromList(KEGG_list), 
      nintersects = 40, 
      nsets = 8, 
      order.by = "freq", 
      decreasing = T, 
      mb.ratio = c(0.6, 0.4),
      number.angles = 0, 
      text.scale = 1.1, 
      point.size = 2.8, 
      line.size = 1
)

# pdf("figures/KEGG_upset.pdf", width = 9, height = 7)
# KEGG_upset
# dev.off()
```

[Figure 60](figures/KEGG_upset.pdf)

```{r fig.width=9, fig.height=7}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 60. Intersections between KEGG pathways."
KEGG_upset
```

### Reactome
```{r}
#| include: false
Reactome_list <- list(PD_ALL = gse.Reactome.df$Description[gse.Reactome.df$NES > 0],
                      Healthy_ALL = gse.Reactome.df$Description[gse.Reactome.df$NES < 0],
                      PD_LRKK2 = gse.Reactome.lrrk2.df$Description[gse.Reactome.lrrk2.df$NES > 0],
                      Healthy_LRKK2 = gse.Reactome.lrrk2.df$Description[gse.Reactome.lrrk2.df$NES < 0],
                      PD_Parkin = gse.Reactome.parkin.df$Description[gse.Reactome.parkin.df$NES > 0],
                      Healthy_Parkin = gse.Reactome.parkin.df$Description[gse.Reactome.parkin.df$NES < 0],
                      LRRK2 = gse.Reactome.B.df$Description[gse.Reactome.B.df$NES > 0],
                      Parkin = gse.Reactome.B.df$Description[gse.Reactome.B.df$NES < 0])

Reactome_upset <- upset(fromList(Reactome_list), 
      nintersects = 40, 
      nsets = 8, 
      order.by = "freq", 
      decreasing = T, 
      mb.ratio = c(0.6, 0.4),
      number.angles = 0, 
      text.scale = 1.1, 
      point.size = 2.8, 
      line.size = 1
)

# pdf("figures/Reactome_upset.pdf", width = 9, height = 7)
# Reactome_upset
# dev.off()
```

[Figure 61](figures/Reactome_upset.pdf)

```{r fig.width=9, fig.height=7}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 61. Intersections between Reactome pathways."
Reactome_upset
```

### HALLMARK
```{r}
#| include: false
all_genes <- unique(unlist(HALLMARK_list))

indicator_matrix <- sapply(HALLMARK_list, function(x) as.integer(all_genes %in% x))
rownames(indicator_matrix) <- all_genes

combinations <- apply(indicator_matrix, 1, paste, collapse = "")
intersection_genes <- split(all_genes, combinations)
unique_intersections <- intersection_genes[sapply(intersection_genes, function(x) length(x) > 0)]

get_set_names <- function(binary_str, set_names) {
  indices <- which(strsplit(binary_str, "")[[1]] == "1")
  paste(set_names[indices], collapse = " & ")
}

set_names <- names(HALLMARK_list)
names(unique_intersections) <- sapply(names(unique_intersections), get_set_names, set_names)

HALLMARK_healthy <- unique_intersections$`Healthy_ALL & Healthy_LRKK2 & Healthy_Parkin`

m_gsea_list <- list(GSEA.H, GSEA.H.lrrk2, GSEA.H.parkin)

HALLMARK_plot <- GSEAmultiGP(gsea_list = m_gsea_list,
            geneSetID = HALLMARK_healthy,
            exp_name = c("PD vs Healthy","LRRK2 vs Healthy","Parkin vs Healthy"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1"))
```

```{r fig.width=7, fig.height=5}
#| echo: false
#| message: false
#| warning: false
download_this(
  HALLMARK_plot,
  output_name = "HALLMARK_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.width=7, fig.height=5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 62. Consistent HALLMARK pathway."
HALLMARK_plot
```

### GO
```{r}
#| include: false
all_genes <- unique(unlist(GO_list))

indicator_matrix <- sapply(GO_list, function(x) as.integer(all_genes %in% x))
rownames(indicator_matrix) <- all_genes

combinations <- apply(indicator_matrix, 1, paste, collapse = "")
intersection_genes <- split(all_genes, combinations)
unique_intersections <- intersection_genes[sapply(intersection_genes, function(x) length(x) > 0)]

get_set_names <- function(binary_str, set_names) {
  indices <- which(strsplit(binary_str, "")[[1]] == "1")
  paste(set_names[indices], collapse = " & ")
}

set_names <- names(GO_list)
names(unique_intersections) <- sapply(names(unique_intersections), get_set_names, set_names)

GO_LRRK2_up <- unique_intersections$`PD_LRKK2 & LRRK2`
GO_LRRK2_down <- unique_intersections$`Healthy_LRKK2 & Parkin`

GO_Parkin_up <- unique_intersections$`PD_Parkin & Parkin`
GO_multidirectional <- unique_intersections$`Healthy_LRKK2 & PD_Parkin & Parkin`

ecm_go_id_1 <- gse.GO.lrrk2.df$ID[gse.GO.lrrk2.df$Description == "extracellular matrix organization"]
ecm_go_id_2 <- gse.GO.lrrk2.df$ID[gse.GO.lrrk2.df$Description == "extracellular structure organization"]
ecm_go_id_3 <- gse.GO.lrrk2.df$ID[gse.GO.lrrk2.df$Description == "external encapsulating structure organization"]
ecm_go_id_4 <- gse.GO.lrrk2.df$ID[gse.GO.lrrk2.df$Description == "collagen-containing extracellular matrix"]
ecm_go_id_5 <- gse.GO.lrrk2.df$ID[gse.GO.lrrk2.df$Description == "external encapsulating structure"]
ecm_go_id_6 <- gse.GO.lrrk2.df$ID[gse.GO.lrrk2.df$Description == "extracellular matrix"]

sod_go_id <- gse.GO.lrrk2.df$ID[gse.GO.lrrk2.df$Description == "sensory organ development"]
eod_go_id <- gse.GO.lrrk2.df$ID[gse.GO.lrrk2.df$Description == "embryonic organ development"]

ribosome_go_id <- gse.GO.lrrk2.df$ID[gse.GO.lrrk2.df$Description == "ribosome"]
ct_go_id <- gse.GO.lrrk2.df$ID[gse.GO.lrrk2.df$Description == "cytoplasmic translation"]

psp_go_id <- gse.GO.lrrk2.df$ID[gse.GO.lrrk2.df$Description == "pattern specification process"]

pdm_go_id <- gse.GO.lrrk2.df$ID[gse.GO.lrrk2.df$Description == "postsynaptic density membrane"]
axon_go_id <- gse.GO.lrrk2.df$ID[gse.GO.lrrk2.df$Description == "distal axon"]
synapse_go_id <- gse.GO.lrrk2.df$ID[gse.GO.lrrk2.df$Description == "regulation of synapse organization"]

go_gsea_list <- list(gse.GO.lrrk2, gse.GO.B)

clrs_go_id <- gse.GO.parkin.df$ID[gse.GO.parkin.df$Description == "cytosolic large ribosomal subunit"]

rs_go_id <- gse.GO.parkin.df$ID[gse.GO.parkin.df$Description == "ribosomal subunit"]
scor_go_id <- gse.GO.parkin.df$ID[gse.GO.parkin.df$Description == "structural constituent of ribosome"]
cr_go_id <- gse.GO.parkin.df$ID[gse.GO.parkin.df$Description == "cytosolic ribosome"]
```

```{r}
#| include: false
ECM_plot_1 <- GSEAmultiGP(gsea_list = go_gsea_list,
            geneSetID = ecm_go_id_1,
            exp_name = c("LRRK2 vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[c(2,4)])

ECM_plot_2 <- GSEAmultiGP(gsea_list = go_gsea_list,
            geneSetID = ecm_go_id_2,
            exp_name = c("LRRK2 vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[c(2,4)])

ECM_plot_3 <- GSEAmultiGP(gsea_list = go_gsea_list,
            geneSetID = ecm_go_id_3,
            exp_name = c("LRRK2 vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[c(2,4)])

ECM_plot_4 <- GSEAmultiGP(gsea_list = go_gsea_list,
            geneSetID = ecm_go_id_4,
            exp_name = c("LRRK2 vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[c(2,4)])

ECM_plot_5 <- GSEAmultiGP(gsea_list = go_gsea_list,
            geneSetID = ecm_go_id_5,
            exp_name = c("LRRK2 vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[c(2,4)])

ECM_plot_6 <- GSEAmultiGP(gsea_list = go_gsea_list,
            geneSetID = ecm_go_id_6,
            exp_name = c("LRRK2 vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[c(2,4)])

SOD_plot <- GSEAmultiGP(gsea_list = go_gsea_list,
            geneSetID = sod_go_id,
            exp_name = c("LRRK2 vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[c(2,4)])

EOD_plot <- GSEAmultiGP(gsea_list = go_gsea_list,
            geneSetID = eod_go_id,
            exp_name = c("LRRK2 vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[c(2,4)])

PDM_plot <- GSEAmultiGP(gsea_list = go_gsea_list,
            geneSetID = pdm_go_id,
            exp_name = c("LRRK2 vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[c(2,4)])

axon_plot <- GSEAmultiGP(gsea_list = go_gsea_list,
            geneSetID = axon_go_id,
            exp_name = c("LRRK2 vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 6, name = "Set1")[c(2,4)])

synapse_plot <- GSEAmultiGP(gsea_list = go_gsea_list,
            geneSetID = synapse_go_id,
            exp_name = c("LRRK2 vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 6, name = "Set1")[c(2,4)])

ribosome_plot <- GSEAmultiGP(gsea_list = go_gsea_list,
            geneSetID = ribosome_go_id,
            exp_name = c("LRRK2 vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[c(2,4)])
```

```{r}
#| include: false
clrs_plot <- GSEAmultiGP(gsea_list = c(gse.GO.parkin, gse.GO.B),
            geneSetID = clrs_go_id,
            exp_name = c("Parkin vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 4, name = "Set1")[c(3,4)])
```

```{r}
#| include: false
rs_plot <- GSEAmultiGP(gsea_list = c(gse.GO.lrrk2, gse.GO.parkin, gse.GO.B),
            geneSetID = rs_go_id,
            exp_name = c("LRRK2 vs Healthy", "Parkin vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 4, name = "Set1")[-1])

scor_plot <- GSEAmultiGP(gsea_list = c(gse.GO.lrrk2, gse.GO.parkin, gse.GO.B),
            geneSetID = scor_go_id,
            exp_name = c("LRRK2 vs Healthy", "Parkin vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 4, name = "Set1")[-1])

cr_plot <- GSEAmultiGP(gsea_list = c(gse.GO.lrrk2, gse.GO.parkin, gse.GO.B),
            geneSetID = cr_go_id,
            exp_name = c("LRRK2 vs Healthy", "Parkin vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 4, name = "Set1")[-1])
```

#### Extracellular matrix
```{r fig.width=7, fig.height=5}
#| echo: false
#| message: false
#| warning: false
download_this(
  ECM_plot_1,
  output_name = "ECM_plot_1",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.width=7.5, fig.height=5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 63. Consistent ECM  pathway for LRRK2 - Extracellular matrix organization."
ECM_plot_1
```

```{r fig.width=7, fig.height=5}
#| echo: false
#| message: false
#| warning: false
download_this(
  ECM_plot_2,
  output_name = "ECM_plot_2",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.width=7.5, fig.height=5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 64. Consistent ECM  pathway for LRRK2 - Extracellular structure organization."
ECM_plot_2
```

```{r fig.width=7, fig.height=5}
#| echo: false
#| message: false
#| warning: false
download_this(
  ECM_plot_3,
  output_name = "ECM_plot_3",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.width=7.5, fig.height=5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 65. Consistent ECM  pathway for LRRK2 - External encapsulating structure organization."
ECM_plot_3
```

```{r fig.width=7, fig.height=5}
#| echo: false
#| message: false
#| warning: false
download_this(
  ECM_plot_4,
  output_name = "ECM_plot_4",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.width=7.5, fig.height=5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 66. Consistent ECM  pathway for LRRK2 - Collagen-containing extracellular matrix."
ECM_plot_4
```

```{r fig.width=7, fig.height=5}
#| echo: false
#| message: false
#| warning: false
download_this(
  ECM_plot_5,
  output_name = "ECM_plot_5",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.width=7.5, fig.height=5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 67. Consistent ECM  pathway for LRRK2 - External encapsulating structure."
ECM_plot_5
```

```{r fig.width=7, fig.height=5}
#| echo: false
#| message: false
#| warning: false
download_this(
  ECM_plot_6,
  output_name = "ECM_plot_6",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.width=7.5, fig.height=5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 68. Consistent ECM  pathway for LRRK2 - Extracellular matrix."
ECM_plot_6
```

#### Development
```{r fig.width=7, fig.height=5}
#| echo: false
#| message: false
#| warning: false
download_this(
  SOD_plot,
  output_name = "SOD_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.width=7.5, fig.height=5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 69. Consistent development pathway for LRRK2 - Sensory organ development."
SOD_plot
```

```{r fig.width=7, fig.height=5}
#| echo: false
#| message: false
#| warning: false
download_this(
  EOD_plot,
  output_name = "SOD_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.width=7.5, fig.height=5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 70. Consistent development pathway for LRRK2 - Embryonic organ development."
EOD_plot
```

#### Synapse
```{r fig.width=7, fig.height=5}
#| echo: false
#| message: false
#| warning: false
download_this(
  PDM_plot,
  output_name = "PDM_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.width=7.5, fig.height=5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 71. Consistent Postsynaptic density membrane pathway for LRRK2."
PDM_plot
```

```{r fig.width=7, fig.height=5}
#| echo: false
#| message: false
#| warning: false
download_this(
  axon_plot,
  output_name = "PDM_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.width=7.5, fig.height=5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 72. Consistent Distal axon pathway for LRRK2."
axon_plot
```

```{r fig.width=7, fig.height=5}
#| echo: false
#| message: false
#| warning: false
download_this(
  synapse_plot,
  output_name = "synapse_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.width=7.5, fig.height=5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 73. Consistent Regulation of synapse organization pathway for LRRK2."
synapse_plot
```

#### Translation
```{r fig.width=7, fig.height=5}
#| echo: false
#| message: false
#| warning: false
download_this(
  ribosome_plot,
  output_name = "ribosome_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.width=7.5, fig.height=5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 74. Ribosome GO pathway."
ribosome_plot
```

```{r fig.width=7, fig.height=5}
#| echo: false
#| message: false
#| warning: false
download_this(
  clrs_plot,
  output_name = "clrs_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.width=7.5, fig.height=5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 75. Cytosolic large ribosomal subunit GO pathway."
clrs_plot
```

```{r fig.width=7, fig.height=5}
#| echo: false
#| message: false
#| warning: false
download_this(
  rs_plot,
  output_name = "rs_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.width=7.5, fig.height=5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 76. Ribosomal subunit GO pathway."
rs_plot
```

```{r fig.width=7, fig.height=5}
#| echo: false
#| message: false
#| warning: false
download_this(
  scor_plot,
  output_name = "scor_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.width=7.5, fig.height=5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 77. Structural constituent of ribosome GO pathway."
scor_plot
```

```{r fig.width=7, fig.height=5}
#| echo: false
#| message: false
#| warning: false
download_this(
  cr_plot,
  output_name = "cr_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 78. Cytosolic ribosome GO pathway."
cr_plot
```

```{r}
#| include: false
all_genes_1 <- unique(unlist(Reactome_list))

indicator_matrix_1 <- sapply(Reactome_list, function(x) as.integer(all_genes_1 %in% x))
rownames(indicator_matrix_1) <- all_genes_1

combinations_1 <- apply(indicator_matrix_1, 1, paste, collapse = "")
intersection_genes_1 <- split(all_genes_1, combinations_1)
unique_intersections_1 <- intersection_genes_1[sapply(intersection_genes_1, function(x) length(x) > 0)]

get_set_names <- function(binary_str, set_names) {
  indices <- which(strsplit(binary_str, "")[[1]] == "1")
  paste(set_names[indices], collapse = " & ")
}

set_names_1 <- names(Reactome_list)
names(unique_intersections_1) <- sapply(names(unique_intersections_1), get_set_names, set_names_1)
```

```{r}
#| include: false
PD <- unique_intersections_1$`PD_ALL & PD_LRKK2 & PD_Parkin`
Healthy <- unique_intersections_1$`Healthy_ALL & Healthy_LRKK2 & Healthy_Parkin`

Parkin_up <- unique_intersections_1$`PD_Parkin & Parkin`
Parkin_down <- unique_intersections_1$`Healthy_Parkin & LRRK2`

LRRK2_up <- unique_intersections_1$`PD_LRKK2 & LRRK2`
LRRK2_down <- unique_intersections_1$`Healthy_LRKK2 & Parkin`

reactome_multi <- unique_intersections_1$`Healthy_LRKK2 & PD_Parkin & Parkin`

tacs_react_id <- gse.Reactome.df$ID[gse.Reactome.df$Description == "Transmission across Chemical Synapses"]

hdr_react_id <- gse.Reactome.df$ID[gse.Reactome.df$Description == "HDR through Homologous Recombination (HRR)"]
ссс_react_id <- gse.Reactome.df$ID[gse.Reactome.df$Description == "Cell Cycle Checkpoints"]

crts_react_id <- gse.Reactome.parkin.df$ID[gse.Reactome.parkin.df$Description == "Cellular response to starvation"]
cd_react_id <- gse.Reactome.parkin.df$ID[gse.Reactome.parkin.df$Description == "Collagen degradation"]
sars1_react_id <- gse.Reactome.parkin.df$ID[gse.Reactome.parkin.df$Description == "SARS-CoV-1 modulates host translation machinery"]
sars2_react_id <- gse.Reactome.parkin.df$ID[gse.Reactome.parkin.df$Description == "SARS-CoV-2 modulates host translation machinery"]

eroge_react_id <- gse.Reactome.parkin.df$ID[gse.Reactome.parkin.df$Description == "Epigenetic regulation of gene expression"]
roer_react_id <- gse.Reactome.parkin.df$ID[gse.Reactome.parkin.df$Description == "Regulation of endogenous retroelements"]
nst_react_id <- gse.Reactome.parkin.df$ID[gse.Reactome.parkin.df$Description == "NGF-stimulated transcription"]

rlr_react_id <- gse.Reactome.lrrk2.df$ID[gse.Reactome.lrrk2.df$Description == "Class A/1 (Rhodopsin-like receptors)"]

rp_react_id <- gse.Reactome.lrrk2.df$ID[gse.Reactome.lrrk2.df$Description == "rRNA processing"]
trans_react_id <- gse.Reactome.lrrk2.df$ID[gse.Reactome.lrrk2.df$Description == "Translation"]
rpnuc_react_id <- gse.Reactome.lrrk2.df$ID[gse.Reactome.lrrk2.df$Description == "rRNA processing in the nucleus and cytosol"]
S43_react_id <- gse.Reactome.lrrk2.df$ID[gse.Reactome.lrrk2.df$Description == "Formation of the ternary complex, and subsequently, the 43S complex"]

aacids_react_id <- gse.Reactome.lrrk2.df$ID[gse.Reactome.lrrk2.df$Description == "Metabolism of amino acids and derivatives"]
influenza_react_id <- gse.Reactome.lrrk2.df$ID[gse.Reactome.lrrk2.df$Description == "Influenza Infection"]
robo_react_id <- gse.Reactome.lrrk2.df$ID[gse.Reactome.lrrk2.df$Description == "Regulation of expression of SLITs and ROBOs"]
GCN2_react_id <- gse.Reactome.lrrk2.df$ID[gse.Reactome.lrrk2.df$Description == "Response of EIF2AK4 (GCN2) to amino acid deficiency"]
nmd1_react_id <- gse.Reactome.lrrk2.df$ID[gse.Reactome.lrrk2.df$Description == "Nonsense-Mediated Decay (NMD)"]
nmd2_react_id <- gse.Reactome.lrrk2.df$ID[gse.Reactome.lrrk2.df$Description == "Nonsense Mediated Decay (NMD) enhanced by the Exon Junction Complex (EJC)"]
srp_react_id <- gse.Reactome.lrrk2.df$ID[gse.Reactome.lrrk2.df$Description == "SRP-dependent cotranslational protein targeting to membrane"]
eti_react_id <- gse.Reactome.lrrk2.df$ID[gse.Reactome.lrrk2.df$Description == "Eukaryotic Translation Initiation"]
cap_react_id <- gse.Reactome.lrrk2.df$ID[gse.Reactome.lrrk2.df$Description == "Cap-dependent Translation Initiation"]
sam_react_id <- gse.Reactome.lrrk2.df$ID[gse.Reactome.lrrk2.df$Description == "Selenoamino acid metabolism"]
viral_react_id <- gse.Reactome.lrrk2.df$ID[gse.Reactome.lrrk2.df$Description == "Influenza Viral RNA Transcription and Replication"]
nmd3_react_id <- gse.Reactome.lrrk2.df$ID[gse.Reactome.lrrk2.df$Description == "Nonsense Mediated Decay (NMD) independent of the Exon Junction Complex (EJC)"]
ett_react_id <- gse.Reactome.lrrk2.df$ID[gse.Reactome.lrrk2.df$Description == "Eukaryotic Translation Termination"]
S60_react_id <- gse.Reactome.lrrk2.df$ID[gse.Reactome.lrrk2.df$Description == "GTP hydrolysis and joining of the 60S ribosomal subunit"]
L13_react_id <- gse.Reactome.lrrk2.df$ID[gse.Reactome.lrrk2.df$Description == "L13a-mediated translational silencing of Ceruloplasmin expression"]
S40_react_id <- gse.Reactome.lrrk2.df$ID[gse.Reactome.lrrk2.df$Description == "Formation of a pool of free 40S subunits"]
```

### Reactome
```{r}
#| include: false
tacs_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome, gse.Reactome.lrrk2, gse.Reactome.parkin),
            geneSetID = tacs_react_id,
            exp_name = c("PD vs Healthy", "LRRK2 vs Healthy","Parkin vs Healthy"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1"))

hdr_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome, gse.Reactome.lrrk2, gse.Reactome.parkin),
            geneSetID = hdr_react_id,
            exp_name = c("PD vs Healthy", "LRRK2 vs Healthy","Parkin vs Healthy"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1"))

ccc_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome, gse.Reactome.lrrk2, gse.Reactome.parkin),
            geneSetID = ссс_react_id,
            exp_name = c("PD vs Healthy", "LRRK2 vs Healthy","Parkin vs Healthy"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1"))
```

```{r}
#| include: false
crts_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.parkin, gse.Reactome.B),
            geneSetID = crts_react_id,
            exp_name = c("Parkin vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[c(3,4)])

cd_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.parkin, gse.Reactome.B),
            geneSetID = cd_react_id,
            exp_name = c("Parkin vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[c(3,4)])

sars1_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.parkin, gse.Reactome.B),
            geneSetID = sars1_react_id,
            exp_name = c("Parkin vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[c(3,4)])

sars2_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.parkin, gse.Reactome.B),
            geneSetID = sars2_react_id,
            exp_name = c("Parkin vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[c(3,4)])
```

```{r}
#| include: false
eroge_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.parkin, gse.Reactome.B),
            geneSetID = eroge_react_id,
            exp_name = c("Parkin vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[c(3,4)])

roer_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.parkin, gse.Reactome.B),
            geneSetID = roer_react_id,
            exp_name = c("Parkin vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[c(3,4)])

nst_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.parkin, gse.Reactome.B),
            geneSetID = nst_react_id,
            exp_name = c("Parkin vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[c(3,4)])
```

```{r}
#| include: false
rlr_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.lrrk2, gse.Reactome.B),
            geneSetID = rlr_react_id,
            exp_name = c("LRRK2 vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[c(2,4)])

rp_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.lrrk2, gse.Reactome.B),
            geneSetID = rp_react_id,
            exp_name = c("LRRK2 vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[c(2,4)])

trans_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.lrrk2, gse.Reactome.B),
            geneSetID = trans_react_id,
            exp_name = c("LRRK2 vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[c(2,4)])

rpnuc_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.lrrk2, gse.Reactome.B),
            geneSetID = rpnuc_react_id,
            exp_name = c("LRRK2 vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[c(2,4)])

S43_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.lrrk2, gse.Reactome.B),
            geneSetID = S43_react_id,
            exp_name = c("LRRK2 vs Healthy","LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[c(2,4)])
```

```{r}
#| include: false
aacids_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.lrrk2, gse.Reactome.parkin, gse.Reactome.B),
            geneSetID = aacids_react_id,
            exp_name = c("LRRK2 vs Healthy", "Parkin vs Healthy", "LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[-1])

influenza_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.lrrk2, gse.Reactome.parkin, gse.Reactome.B),
            geneSetID = influenza_react_id,
            exp_name = c("LRRK2 vs Healthy", "Parkin vs Healthy", "LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[-1])

robo_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.lrrk2, gse.Reactome.parkin, gse.Reactome.B),
            geneSetID = robo_react_id,
            exp_name = c("LRRK2 vs Healthy", "Parkin vs Healthy", "LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[-1])

GCN2_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.lrrk2, gse.Reactome.parkin, gse.Reactome.B),
            geneSetID = GCN2_react_id,
            exp_name = c("LRRK2 vs Healthy", "Parkin vs Healthy", "LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[-1])

nmd1_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.lrrk2, gse.Reactome.parkin, gse.Reactome.B),
            geneSetID = nmd1_react_id,
            exp_name = c("LRRK2 vs Healthy", "Parkin vs Healthy", "LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[-1])

nmd2_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.lrrk2, gse.Reactome.parkin, gse.Reactome.B),
            geneSetID = nmd2_react_id,
            exp_name = c("LRRK2 vs Healthy", "Parkin vs Healthy", "LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[-1])

nmd3_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.lrrk2, gse.Reactome.parkin, gse.Reactome.B),
            geneSetID = nmd3_react_id,
            exp_name = c("LRRK2 vs Healthy", "Parkin vs Healthy", "LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[-1])

srp_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.lrrk2, gse.Reactome.parkin, gse.Reactome.B),
            geneSetID = srp_react_id,
            exp_name = c("LRRK2 vs Healthy", "Parkin vs Healthy", "LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[-1])

eti_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.lrrk2, gse.Reactome.parkin, gse.Reactome.B),
            geneSetID = eti_react_id,
            exp_name = c("LRRK2 vs Healthy", "Parkin vs Healthy", "LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[-1])

cap_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.lrrk2, gse.Reactome.parkin, gse.Reactome.B),
            geneSetID = cap_react_id,
            exp_name = c("LRRK2 vs Healthy", "Parkin vs Healthy", "LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[-1])

sam_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.lrrk2, gse.Reactome.parkin, gse.Reactome.B),
            geneSetID = sam_react_id,
            exp_name = c("LRRK2 vs Healthy", "Parkin vs Healthy", "LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[-1])

viral_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.lrrk2, gse.Reactome.parkin, gse.Reactome.B),
            geneSetID = viral_react_id,
            exp_name = c("LRRK2 vs Healthy", "Parkin vs Healthy", "LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[-1])

ett_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.lrrk2, gse.Reactome.parkin, gse.Reactome.B),
            geneSetID = ett_react_id,
            exp_name = c("LRRK2 vs Healthy", "Parkin vs Healthy", "LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[-1])

S60_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.lrrk2, gse.Reactome.parkin, gse.Reactome.B),
            geneSetID = S60_react_id,
            exp_name = c("LRRK2 vs Healthy", "Parkin vs Healthy", "LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[-1])

L13_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.lrrk2, gse.Reactome.parkin, gse.Reactome.B),
            geneSetID = L13_react_id,
            exp_name = c("LRRK2 vs Healthy", "Parkin vs Healthy", "LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[-1])

S40_plot <- GSEAmultiGP(gsea_list = c(gse.Reactome.lrrk2, gse.Reactome.parkin, gse.Reactome.B),
            geneSetID = S40_react_id,
            exp_name = c("LRRK2 vs Healthy", "Parkin vs Healthy", "LRRK2 vs Parkin"), 
            legend.position = "right", 
            curve.col = brewer.pal(n = 5, name = "Set1")[-1])
```

#### Amino acid metabolism and integration with translation
```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  aacids_plot,
  output_name = "aacids_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 79. Metabolism of amino acids and derivatives Reactome pathway."
aacids_plot
```

```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  sam_plot,
  output_name = "sam_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 80. Selenoamino acid metabolism Reactome pathway."
sam_plot
```

```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  GCN2_plot,
  output_name = "GCN2_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 81. Response of EIF2AK4 (GCN2) to amino acid deficiency Reactome pathway."
GCN2_plot
```

#### Transcription and epigenetic regulation
```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  eroge_plot,
  output_name = "eroge_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 82. Epigenetic regulation of gene expression."
eroge_plot
```

```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  roer_plot,
  output_name = "roer_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 83. Regulation of endogenous retroelements."
roer_plot
```

```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  nst_plot,
  output_name = "nst_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 84. NGF-stimulated transcription."
nst_plot
```

```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  robo_plot,
  output_name = "robo_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 85. Regulation of expression of SLITs and ROBOs."
robo_plot
```

#### Protein Synthesis: Initiation, Elongation, Termination
```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  rp_plot,
  output_name = "rp_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 86. rRNA processing."
rp_plot
```

```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  S40_plot,
  output_name = "S40_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 87. Formation of a pool of free 40S subunits."
S40_plot
```

```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  S43_plot,
  output_name = "S43_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 88. Formation of the ternary complex, and subsequently, the 43S complex."
S43_plot
```

```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  cap_plot,
  output_name = "S43_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 89. Cap-dependent Translation Initiation."
cap_plot
```

```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  S60_plot,
  output_name = "S60_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 90. GTP hydrolysis and joining of the 60S ribosomal subunit."
S60_plot
```

```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  trans_plot,
  output_name = "trans_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 89. Translation."
trans_plot
```

```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  ett_plot,
  output_name = "ett_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 90. Eukaryotic Translation Termination."
ett_plot
```

#### Post-translational events and quality control
```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  srp_plot,
  output_name = "srp_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 91. SRP-dependent cotranslational protein targeting to membrane."
srp_plot
```

```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  nmd1_plot,
  output_name = "nmd1_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 92. Nonsense mediated decay."
nmd1_plot
```

```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  L13_plot,
  output_name = "L13_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 93. L13a-mediated translational silencing."
L13_plot
```

#### Cellular communication and signaling
```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  rlr_plot,
  output_name = "rlr_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 94. Class A/1 (Rhodopsin-like receptors)."
rlr_plot
```

```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  tacs_plot,
  output_name = "tacs_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 95. Transmission across Chemical Synapses."
tacs_plot
```

#### Cellular response to stress
```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  crts_plot,
  output_name = "crts_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 96. Cellular response to starvation."
crts_plot
```

#### Virus-host interaction
```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  influenza_plot,
  output_name = "influenza_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 97. Influenza Infection."
influenza_plot
```

```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  viral_plot,
  output_name = "viral_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 98. Influenza Viral RNA Transcription and Replication."
viral_plot
```

```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  sars1_plot,
  output_name = "sars1_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 99. SARS-CoV-1 modulates host translation machinery."
sars1_plot
```

```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  sars2_plot,
  output_name = "sars2_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 100. SARS-CoV-2 modulates host translation machinery."
sars2_plot
```

#### DNA repair and cell cycle control
```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  hdr_plot,
  output_name = "hdr_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 101. HDR through Homologous Recombination (HRR)."
hdr_plot
```

```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  ccc_plot,
  output_name = "ccc_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 102. Cell Cycle Checkpoints."
ccc_plot
```

#### Extracellular matrix and its remodeling
```{r fig.height=5, fig.width=7}
#| echo: false
#| message: false
#| warning: false
download_this(
  cd_plot,
  output_name = "cd_plot",
  output_extension = ".pdf",
  button_label = "PDF",
  button_type = "primary"
)
```

```{r fig.height=5, fig.width=7.5}
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Figure 103. Collagen degradation."
cd_plot
```

## Table 77. Comparison of GSEA results in Parkinson's disease models (LRRK2 vs. Parkin)
| Feature | LRRK2 | Parkin | Biological Interpretation |
| :--- | :--- | :--- | :--- |
| **Synaptic Transmission** | ↑ Transmission across Chemical Synapses | ↑ Transmission across Chemical Synapses | **Common Mechanism:** |
| **Cell Cycle & DNA Repair** | ↓ E2F Targets, ↓ Cell Cycle Checkpoints, ↓ HRR | ↓ E2F Targets, ↓ Cell Cycle Checkpoints, ↓ HRR | **Common Mechanism:**| 
| **Protein Translation** | ↓ Ribosomal subunits, ↓ Translation initiation | ↑ Ribosomal subunits, ↑ Translation initiation | **Antiphase:** |
| **Viral Response** | ↓ Influenza infection pathways | ↑ SARS-CoV-1/2, ↑ Influenza infection pathways | **Antiphase:** |
| **Synaptic Structure** | ↑ Postsynaptic density, ↑ Distal axon, ↑ Synapse organization | — | **LRRK2-Unique:** |
| **Extracellular Matrix** | ↓ All major ECM pathways | — | **LRRK2-Unique:**|
| **Amino Acid Metabolism** | — | ↑ Metabolism of amino acids, ↑ Selenoamino acid metabolism | **Parkin-Unique:** |
| **Epigenetic Regulation** | — | ↓ Epigenetic regulation, ↓ NGF-stimulated transcription | **Parkin-Unique:** |